---
title: "FInding solar markets for investment, using html tables, GIS, and Twitter - an exercise"
author: "Ram A. Barankin"
date: "2/24/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The goal
Demonstrate data acquisition, data preparation, data analysis and visualization.

(For required files: <https://github.com/RamBarankin/Solar_Market.git>).

## Data types
- Local GIS shapefile
- HTML table (scraped from the web)
- Twitter data (API using the "twitteR" package)

## The exercise
Find the best place to invest in current patterns according to local climate, current electricity prices, and attitude toward climate change.

###1. Load general packages
```{r packages}
library(XML)
library(dplyr)
```

###2. Find lat/long of cities (see GIS shapefile in the folder)
```{r LatLong}
library(raster)
cities <- tbl_df(shapefile("Analysis_files//US_cities/US_cities.shp")) 
cities <- cities %>%  plyr::rename(c("ycoord"="LAT")) %>% plyr::rename(c("xcoord"="LONG")) %>% 
  filter(CLASS=="city") %>% dplyr::select(NAME, ST, LAT, LONG)
```
This is the resulted dataframe:
```{r LatLong Table, echo=FALSE}
cities
```

###3. Get average electricity prices
```{r prices}
library(RCurl)
data <- getURL("https://www.eia.gov/electricity/monthly/epm_table_grapher.cfm?t=epmt_5_6_a", ssl.verifypeer = FALSE)
perftable <- readHTMLTable(data[[1]], stringsAsFactors = T)
elctprice <- perftable[[2]]
elctprice <- tbl_df(elctprice[,c(1,10)])
colnames(elctprice) <- c("State", "Price")
```
This is the resulted dataframe:
```{r prices Table, echo=FALSE}
elctprice
```

###4. Add State's abbreviation (to have a common column for the two dataframes)
```{r abb}
us <- readHTMLTable("http://www.softschools.com/social_studies/state_abbreviations/",  
                    header=T, which=1,stringsAsFactors=F, trim = T, skip.rows = 3)
colnames(us) <- us[3,]
us <- tbl_df(us[4:nrow(us),])
elctprice <- elctprice %>% mutate(State=toupper(State)) %>%
  left_join(us) %>% dplyr::select(Abbreviation, Price) %>% plyr::rename(c("Abbreviation"="ST")) %>% na.omit()
```
This is the resulted dataframe:
```{r prices Table2, echo=FALSE}
elctprice
```

###5. Merge all and produce the index
```{r merge}
cities <- cities %>% left_join(elctprice)
cities$Price <- as.numeric(cities$Price)
cities <- cities %>% na.omit() %>% mutate(idx_SUN = scale(LAT), idx_SUN=idx_SUN+abs(min(idx_SUN)),
                            idx_PRICE = scale(Price), idx_PRICE=idx_PRICE+abs(min(idx_PRICE)),
                            Market_Potential_Index = idx_SUN+idx_PRICE)
Market_Potential <- quantile(cities$Market_Potential_Index)
for (i in 1:nrow(cities)){
if (cities$Market_Potential_Index[i]<=Market_Potential[2]){cities$Market_Potential[i]<-"High"} else{
  if(cities$Market_Potential_Index[i]>Market_Potential[2]&cities$Market_Potential_Index[i]<=Market_Potential[3]){ cities$Market_Potential[i]<-"Medium"} else{
    cities$Market_Potential[i]<-"Low"}}}
  
cities$Market_Potential_Index <- -cities$Market_Potential_Index+9
```
This is the resulted dataframe:
```{r Index, echo=FALSE}
cities
```
###6. Map
```{r map}
df <- cities
df <- df%>% plyr::rename(c("Market_Potential_Index"="Solar Market Potential index"))%>%
  plyr::rename(c("Market_Potential"="Market Potential Level")) %>% 
  plyr::rename(c("Price"="Electricity Price (Cents per KW)"))
map <- shapefile("Analysis_files/US_cities/US_cities.shp")
map <- merge(map, df)

map1 <- map[,c(1,2, 54, 57,58)]
map1 <- map1[na.omit(map1),]
library(mapview)
m <- mapview(map1, legend=T, zcol="Solar Market Potential index")
```
The map:
```{r map1, echo=F}
m
```

###7. Discover attitudes using twitter
```{r twit, eval = FALSE}
library(twitteR)
# setup_twitter_oauth (use yourown credentials)
cities <- cities[order(-cities$Market_Potential_Index),]
poten_cities <- cities[1:10,] # The cities with highest potential
# prepare Twitter parameters:
poten_cities$LatLong <- paste0(poten_cities$LAT, ",", poten_cities$LONG,
                               ", ", " 2mi") #location
date = as.character(Sys.Date()) # Today's date
poten_cities$Twitnum <- 0 # the variable to be used to count tweats
for (i in 1:nrow(poten_cities)){
  tw = searchTwitter('#climatechange', n = 1e4, since = date, geocode=poten_cities$LatLong[i])
  poten_cities$Twitnum[i] <- length(tw)
  } #get and store the # of tweats
poten_cities$Potential <- scale(poten_cities$Twitnum) #standardize the # of tweats 
poten_cities <- poten_cities[order(-poten_cities$Potential),]#order cities according to tweets
```
The resulting dataframe:
```{r tweattable, echo=F}
load('Analysis_files/potencities.rda')
poten_cities
```

###8. Visualize 
```{r tweatgraph1}
library(plotly)
poten_cities$NAME <- factor(poten_cities$NAME, levels = poten_cities[["NAME"]])

p <- plot_ly(poten_cities, x=~poten_cities$NAME ,y=~poten_cities$Potential, type="bar", 
        marker = list(color = 'rgb(158,202,225)', range ("-1", "-0.5", "Mean", "0.5", "1"),
                      line = list(color = 'rgb(8,48,107)', width = 1.5))) %>%
  layout(title = "Solar Market Potential - Twitter Analysis",
         xaxis = list(title = ""),
         yaxis = list(title = "#climatechange Tweets Relative to the Mean"))
       
```      
The graph:
```{r tweatgraph, echo=F}
p
```

